Unshuffling Permutations

### 2015-12-03 submission.tex
Petites modifs

As announced, let us now use $\Delta$ to define a shuffle product on
permutations. As any coproduct, $\Delta$ leads to the definition of a
product obtained by duality in the following way.
From~\eqref{equ:unshuffling_coproduct}, for any permutation $\pi$, we
have
\begin{equation} \label{equ:schematic_coproduct}
    \Delta(\pi) =
    \sum_{\sigma, \nu \in S} \lambda_{\sigma, \nu}^\pi \;
    \sigma \otimes \nu,
\end{equation}
where the $\lambda_{\sigma, \nu}^\pi$ are nonnegative integers. Now,
by definition of duality, the dual product of $\Delta$, denoted by
$\SHUFFLE$, is a linear binary product on $\QQ[S]$. It satisfies, for
any permutations $\sigma$ and $\nu$,
\begin{equation}
    \sigma \SHUFFLE \nu =
    \sum_{\pi \in S}
    \lambda_{\sigma, \nu}^\pi \; \pi,
\end{equation}
where the coefficients $\lambda_{\sigma, \nu}^\pi$ are the ones
of~\eqref{equ:schematic_coproduct}. We call $\SHUFFLE$ the
{\em shuffle product of permutations}. For instance,
\begin{equation}\begin{split} \label{equ:example_shuffle_product}
    12 \SHUFFLE 21 & =
    1243 + 1324 + 2.1342 + 2.1423 + 3. {\bf 1432} + 2134 + 2.2314 \\
    & + 3.2341 + 2413 + 2.2431 + 2.3124 + 3142 + 3.3214 + 2.3241 \\
    & + 3421 + 3.4123 + 2.4132 + 2.4213 + 4231 + 4312.
\end{split}\end{equation}
Observe that the coefficient $3$ of the permutation $1432$
in~\eqref{equ:example_shuffle_product} comes from the fact that the
coefficient of the tensor $12 \otimes 21$ is $3$
in~\eqref{equ:example_unshuffling_coproduct}.

>>>>

As announced, let us now use $\Delta$ to define a shuffle product on
permutations. As any coproduct, $\Delta$ leads to the definition of a
product obtained by duality in the following way.
From~\eqref{equ:unshuffling_coproduct}, for any permutation $\pi$, we
have
\begin{equation} \label{equ:schematic_coproduct}
    \Delta(\pi) =
    \sum_{\sigma, \nu \in S} \lambda_{\sigma, \nu}^\pi \;
    \sigma \otimes \nu,
\end{equation}
where the $\lambda_{\sigma, \nu}^\pi$ are nonnegative integers. Now,
by definition of duality, the dual product of $\Delta$, denoted by
$\SHUFFLE$, is a linear binary product on $\QQ[S]$. It satisfies, for
any permutations $\sigma$ and $\nu$,
\begin{equation}
    \sigma \SHUFFLE \nu =
    \sum_{\pi \in S}
    \lambda_{\sigma, \nu}^\pi \; \pi,
\end{equation}
where the coefficients $\lambda_{\sigma, \nu}^\pi$ are the ones
of~\eqref{equ:schematic_coproduct}. We call $\SHUFFLE$ the
{\em shuffle product of permutations}. For instance,
\begin{equation}\begin{split} \label{equ:example_shuffle_product}
    12 \SHUFFLE 21 & =
    1243 + 1324 + 2 \cdot 1342 + 2 \cdot 1423 + 3 \cdot {\bf 1432} + 
    2134 + 2 \cdot 2314 \\
    & + 3 \cdot 2341 + 2413 + 2 \cdot 2431 + 2 \cdot 3124 + 3142 + 
    3 \cdot 3214 + 2 \cdot 3241 \\
    & + 3421 + 3 \cdot 4123 + 2 \cdot 4132 + 2 \cdot 4213 + 4231 + 4312.
\end{split}\end{equation}
Observe that the coefficient $3$ of the permutation $1432$
in~\eqref{equ:example_shuffle_product} comes from the fact that the
coefficient of the tensor $12 \otimes 21$ is $3$
in~\eqref{equ:example_unshuffling_coproduct}.
